---
title: "MicroHub Workshop Weekly Report"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format: 
  html:
    theme: yeti
    toc: true
    toc-title: "Jump To"
    embed-resources: true
    smooth-scroll: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: set-up
#| include: false

# load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, here, DT, scoringutils, distfromq, hubEnsembles, reshape2, hubUtils, plotly, rmarkdown, knitr, kableExtra, ISOweek, MMWRweek, htmltools)

# settings
current_date <- as.Date("2026-06-24")  # Sys.Date()
current_ref  <- lubridate::ceiling_date(current_date, "week") - lubridate::days(1)

# forecasting files
forc_files <- list.files(
  path    = here("auxiliary-data", "microhub-output"),
  pattern = ".*-output_\\d{4}-\\d{2}-\\d{2}\\.csv$",
  full.names = TRUE
)
forc_dates <- regmatches(forc_files, regexpr("\\d{4}-\\d{2}-\\d{2}", forc_files)) %>%
  as.Date()
first_forecast <- forc_dates %>%
  # Keep only dates in the same year as current_ref
  .[epiyear(.) == epiyear(current_ref)] %>%
  min()

```

```{r}
#| label: truth-dat
#| include: false

# Truth data this week
raw_truth <- read_csv(
  here("target-data", "chile-target.csv"),
    col_types = cols_only(
      date = col_character(),
      target_group = col_character(),
      value = col_double())
   ) %>% 
  mutate(date = as.Date(parse_date_time(date, orders = c("Ymd", "mdy", "mdY")))) %>% 
  arrange(date) 

```


```{r}
#| label: fig1-ensemble-prep
#| include: false

# Helper function: convert quantile to pmf
quantile_to_pmf <- function(model_id = "Copycat",
                            current_forecasts = forecast_tbl,
                            threshold = diff_data, 
                            categories = c("large_increase", "increase", "stable", "decrease", "large_decrease"), 
                            horizons = 0:4) {
  
  num_cat = length(categories)
  
  quantile_forecasts <- current_forecasts %>% filter(model == model_id) %>% mutate(output_type_id = as.numeric(output_type_id))
  
  pmf_ready <- quantile_forecasts %>% 
    left_join(threshold %>% 
                mutate(crit1 = value + crit1,
                       crit2 = value + crit2,
                       crit3 = value + crit3,
                       crit4 = value + crit4) %>% 
                mutate(horizon = wk_diff+(horizons[1]-1)) %>% 
                ungroup() %>% 
                select(-wk_diff, -value),
              by = c("target_group", "horizon")) 
  
  train_forecasts <- pmf_ready %>%
    dplyr::distinct(model, target_group, reference_date, horizon, .keep_all=TRUE) %>%
    dplyr::select(-output_type,-output_type_id,-value)
  
  for (i in 1:(num_cat-1)) {
    pmf_ready[["crit_current"]] <- pmf_ready[[paste0("crit", i, sep="")]] 
    temp <- pmf_ready %>%
      dplyr::group_by(model, reference_date, target_group, horizon) %>%
      dplyr::summarize(
        cdf_crit_current = distfromq::make_p_fn(
          ps = output_type_id,
          qs = value)(unique(crit_current), log = FALSE)
      ) %>%
      dplyr::ungroup() 
    train_forecasts[[paste0("cdf_crit", i)]] <- temp[["cdf_crit_current"]]
  }
  
  #calculate category percentages from cdf criteria, correcting for negative numbers
  exp_forecast <- train_forecasts %>%
    dplyr::ungroup() %>%
    dplyr::mutate(cdf_crit0=1, .before=cdf_crit1) 
  
  exp_forecast[[paste0("cdf_crit", num_cat)]] <- 0
  
  exp_forecast <- exp_forecast %>%
    mutate(!!categories[1] := cdf_crit0 - cdf_crit1)
  
  # Initialize cumulative sum of crits starting with the first category
  cdf_crit_sum <- sym(categories[1])
  
  # Generalizable loop for the rest of the categories
  for (i in 2:length(categories)) {
    exp_forecast <- exp_forecast %>%
      mutate(!!categories[i] := ifelse(!!cdf_crit_sum < 1, 
                                       get(paste0("cdf_crit", i - 1)) - get(paste0("cdf_crit", i)), 
                                       0))
    
    # Update cumulative sum with the newly created column
    cdf_crit_sum <- quo(!!cdf_crit_sum + !!sym(categories[i]))
  }
  
  exp_forecast <- exp_forecast %>%
    dplyr::select(model, reference_date, target_group, horizon, all_of(categories))
  
  
  #transpose data_frame to format for submission
  exp_t = reshape2::melt(
    exp_forecast,
    id.vars = c("model", "reference_date", "target_group","horizon"),
    measure.vars = categories,
    variable.name="output_type_id",
    value.name="value"
  )
  
  exp_t <- exp_t %>%
    dplyr::mutate(output_type="pmf") %>%
    dplyr::select(model, reference_date, horizon, target_group, output_type, output_type_id, value)
  
  output_forecasts <- exp_t %>%
    dplyr::mutate(
      output_type_id=as.character(output_type_id),
      target_end_date=reference_date+weeks(horizon)
    )
  
  return (output_forecasts)
}


```

## Current week's forecast: new SARI cases

The figure shows the number of severe acute respiratory infections (SARI) cases reported through FluID. The black line represents reported cases. Forecasts include estimates for the previous week (hindcast, in purple), the current week (nowcast, in dark blue), and the 3-week ahead forecasts in the blue line with corresponding prediction intervals.

Ensemble forecasts combine forecasts from different models into one forecast. These ensemble forecasts are generated using an ensemble of three previously validated models: models that focus on historical patterns in the data (smoothing and pattern matching approaches) and a model based on a machine learning method (gradient boosting).

Each week is also summarized as one of five trend categories, shown in the colored strip below the figure, based on the direction and magnitude of change in case counts – large increase, increase, stable, decrease, large decrease. For reported weeks, the trend compares each week with the week immediately before it (a week-to-week change). For forecasted weeks, the trend instead compares each week with the most recent reported week; because that reference point stays fixed, the change is measured over a longer window the further ahead the forecast reaches (a one-week change for the first forecasted week, a two-week change for the next, and so on).The thresholds for these categories were determined using data from 2022, or the earliest year available, through the previous year (see appendix for details). 

```{r}
#| label: fig1-check
#| include: false

forc_filepath <- paste0(here("auxiliary-data", "microhub-output"),
                         "/microhub-output_", current_ref, ".csv")
forc_available <- file.exists(forc_filepath)
```

```{r}
#| label: fig1-forc-prep
#| include: false
#| eval: !expr forc_available
 
# ensemble forecast output for this week
forecast <- readr::read_csv(
  forc_filepath,
  col_types = cols_only(
    model = col_character(),
    reference_date = col_character(),
    horizon = col_integer(),
    target_end_date = col_character(),
    target_group = col_character(),
    output_type = col_character(),
    output_type_id = col_character(),
    value = col_double()
    )
  ) %>% 
  mutate(reference_date = as.Date(parse_date_time(reference_date, orders = c("Ymd", "mdy", "mdY"))),
         target_end_date = as.Date(parse_date_time(target_end_date, orders = c("Ymd", "mdy", "mdY")))) %>% 
  filter(output_type == "quantile") 

# plotly data to work with
common_dates <- intersect(raw_truth %>% pull(date), 
                          forecast$target_end_date)
fig1_dat <- bind_rows(raw_truth %>% 
                        filter(date < min(forecast$target_end_date)) %>% 
                        dplyr::select(target_end_date = date, target_group, value),
                      forecast %>% 
                        filter(model == "Ensemble") %>%
                        pivot_wider(names_from = output_type_id,
                                    values_from = value) %>% 
                        dplyr::select(forc_value = `0.5`, target_end_date, target_group, 
                             `0.025`, `0.975`, `0.25`, `0.75`)) %>% 
  filter(target_end_date > as.Date(current_ref) - 150) %>% 
  filter(target_end_date < as.Date(current_ref) + 30) %>% 
  group_by(target_group) %>% 
  complete(target_end_date = seq(min(target_end_date), max(target_end_date), by = "1 week")) %>% 
  mutate(
    hindcast_drop = case_when(target_end_date %in% common_dates ~ forc_value, TRUE ~ NA_real_),
    hindcast = case_when(!(target_end_date %in% common_dates) & 
                           target_end_date < current_ref ~ forc_value, TRUE ~ NA_real_),
    nowcast  = case_when(target_end_date == current_ref ~ forc_value, TRUE ~ NA_real_),
    forecast = case_when(target_end_date > current_ref ~ forc_value, TRUE ~ NA_real_)
  ) %>% 
  mutate(
    week_label = if_else(
      epiweek(target_end_date) == 1,
      paste0(lubridate::epiyear(target_end_date), " Week ", epiweek(target_end_date)),
      paste0("Week ", epiweek(target_end_date)))
    ) 
  
# Create a list of unique dates and their corresponding epiweek labels for tickvals and ticktext
tickvals <- unique(fig1_dat$target_end_date)
ticktext <- unique(fig1_dat$week_label)

# x axis tick
two_weeks_ms <- 14 * 24 * 60 * 60 * 1000

# Split the data by 'group' and create a plot for each subset
fig1 <- lapply(unique(fig1_dat$target_group), function(g) {
  
  dat <- fig1_dat %>% filter(target_group == g) %>% 
    mutate(
      purple_line = coalesce(hindcast_drop, hindcast, nowcast),
      blue_line = coalesce(nowcast, forecast)
    )
  
  p <- plot_ly(data = dat, 
               x = ~target_end_date) %>% 
  add_trace(y = ~`0.025`, name = "outer_lower", 
            type = "scatter", mode = "lines", line = list(color = "transparent"),
            showlegend = FALSE,
            yaxis = 'y2') %>% 
  add_trace(y = ~`0.975`, name = "outer_upper", 
            type = "scatter", mode = "lines", line = list(color = "transparent"),
            fill = "tonexty", fillcolor="#CBEAFB", 
            showlegend = FALSE,
            yaxis = 'y2') %>% 
  add_trace(y = ~`0.25`, name = "inner_lower", 
            type = "scatter", mode = "lines", line = list(color = "transparent"),
            showlegend = FALSE,
            yaxis = 'y2') %>% 
  add_trace(y = ~`0.75`, name = "inner_up", 
            type = "scatter", mode = "lines", line = list(color = "transparent"),
            fill = "tonexty", fillcolor="#5CC3F0", 
            showlegend = FALSE,
            yaxis = 'y2') %>% 
  add_trace(y = ~value, name = "Reported", 
            type = "scatter", mode = "lines+markers", 
            line = list(color = "black"),
            marker = list(color = "black", size = 8),
            yaxis = 'y2') %>% 
  add_trace(y = ~purple_line, name = "purple_line",
            type = "scatter", mode = "lines",
            line = list(color = "#7B6FC4"),
            showlegend = FALSE, yaxis = 'y2') %>%
  add_trace(y = ~blue_line, name = "blue_line",
            type = "scatter", mode = "lines",
            line = list(color = "#2B55D4"),
            showlegend = FALSE, yaxis = 'y2') %>%
  add_trace(y = ~hindcast_drop, name = "Hindcast (dropped weeks)",
          type = "scatter", mode = "markers",
          # line = list(color = "#7B6FC4"),
          marker = list(color = "transparent", size = 8, line = list(color = "#7B6FC4", width = 2)),
          showlegend = FALSE,
          yaxis = 'y2') %>%
  add_trace(y = ~hindcast, name = "Hindcast",
          type = "scatter", mode = "markers",
          # line = list(color = "#7B6FC4"),
          marker = list(color = "#7B6FC4", size = 8),
          yaxis = 'y2') %>%
  add_trace(y = ~nowcast, name = "Nowcast",
          type = "scatter", mode = "markers",
          # line = list(color = "#1A3A8F"),
          marker = list(color = "#1A3A8F", size = 8),
          yaxis = 'y2') %>%
  add_trace(y = ~forecast, name = "Forecast",
          type = "scatter", mode = "markers",
          # line = list(color = "#2B55D4"),
          marker = list(color = "#2B55D4", size = 8),
          yaxis = 'y2') %>% 
  add_trace(y = ~value, "Y1-axis",
            type = "scatter", mode = "lines",
            line = list(color = "transparent"),
            showlegend = FALSE,
            yaxis = 'y1') %>% 
  layout(
    title = list(text = "", 
                 x = 0.05, 
                 font = list(family="Arial", size = 18, style = "bold")),
    xaxis = list(title = "",
                 type = "date", 
                 tick0 = min(fig1_dat$target_end_date),
                 dtick = two_weeks_ms,
                 range = c(as.Date(min(fig1_dat$target_end_date))-5, as.Date(max(fig1_dat$target_end_date))+5),
                 showgrid = TRUE,          # Disable grid lines on the y-axis
                 showline = FALSE,         # Remove axis line
                 tickvals = tickvals,
                 showticklabels = FALSE    # No tick labels on the y-axis
                ),
    yaxis = list(showticklabels = FALSE, showgrid = FALSE,
                 range = c(0, max(c(dat$value, dat$`0.975`), 
                                  na.rm = TRUE)*1.2)),
    yaxis2 = list(side = "right", showline = TRUE, showticklabels = TRUE, 
                  tickangle = 0, tickfont = list(family="Arial", size = 18),
                  range = c(0, max(c(dat$value, dat$`0.975`), 
                                   na.rm = TRUE)*1.2)),
    legend = list(
      x = 0.01,
      y = 0.99,
      orientation = "v",
      font = list(family = "Arial", size = 12)
    ),
    annotations = list(
      list(x = 0.99, y = 0.98, xref = "paper", yref = "paper", 
           text = "Inner Bands: 50% Prediction Intervals", font = list(family="Arial", size = 12),
           showarrow = FALSE),
      list(x = 0.99, y = 0.93, xref = "paper", yref = "paper", 
           text = "Outer Bands: 95% Prediction Intervals", font = list(family="Arial", size = 12),
           showarrow = FALSE)
      ),
    margin = list(t = 50, r = 50)
  )
  
  return(p)
})

```

```{r}
#| label: fig1-diff-prep
#| include: false
#| eval: !expr forc_available

# most recent week
most_recent <- raw_truth %>% 
  group_by(target_group) %>% 
  filter(date < min(forecast$target_end_date)) %>% 
  slice(n())

# week difference using historical data
# The main use for this is to convert absolute number to historical X-week difference
diff_data <- raw_truth %>% 
  filter(epiyear(date) < year(Sys.Date())) %>%  # we are evaluating up to previous season
  group_by(target_group) %>%
  arrange(date) %>%
  mutate(flu_diff_h0 = value - lag(value, 1),
         flu_diff_h1 = value - lag(value, 2),
         flu_diff_h2 = value - lag(value, 3),
         flu_diff_h3 = value - lag(value, 4),
         flu_diff_h4 = value - lag(value, 5),
         flu_diff_h5 = value - lag(value, 6),
         flu_diff_h6 = value - lag(value, 7),
         flu_diff_h7 = value - lag(value, 8),
         flu_diff_h8 = value - lag(value, 9)) %>%
  summarise(h0_percentile_05 = quantile(flu_diff_h0, probs = .05, na.rm = TRUE), #large decrease
            h0_percentile_25  = quantile(flu_diff_h0, probs = .25,  na.rm = TRUE), #decrease
            h0_percentile_75  = quantile(flu_diff_h0, probs = .75,  na.rm = TRUE), #increase
            h0_percentile_95 = quantile(flu_diff_h0, probs = .95, na.rm = TRUE), #large increase

            h1_percentile_05 = quantile(flu_diff_h1, probs = .05, na.rm = TRUE), #large decrease
            h1_percentile_25  = quantile(flu_diff_h1, probs = .25,  na.rm = TRUE), #decrease
            h1_percentile_75  = quantile(flu_diff_h1, probs = .75,  na.rm = TRUE), #increase
            h1_percentile_95 = quantile(flu_diff_h1, probs = .95, na.rm = TRUE), #large increase

            h2_percentile_05 = quantile(flu_diff_h2, probs = .05, na.rm = TRUE), #large decrease
            h2_percentile_25  = quantile(flu_diff_h2, probs = .25,  na.rm = TRUE), #decrease
            h2_percentile_75  = quantile(flu_diff_h2, probs = .75,  na.rm = TRUE), #increase
            h2_percentile_95 = quantile(flu_diff_h2, probs = .95, na.rm = TRUE), #large increase

            h3_percentile_05 = quantile(flu_diff_h3, probs = .05, na.rm = TRUE), #large decrease
            h3_percentile_25  = quantile(flu_diff_h3, probs = .25,  na.rm = TRUE), #decrease
            h3_percentile_75  = quantile(flu_diff_h3, probs = .75,  na.rm = TRUE), #increase
            h3_percentile_95 = quantile(flu_diff_h3, probs = .95, na.rm = TRUE), #large increase

            h4_percentile_05 = quantile(flu_diff_h4, probs = .05, na.rm = TRUE), #large decrease
            h4_percentile_25  = quantile(flu_diff_h4, probs = .25,  na.rm = TRUE), #decrease
            h4_percentile_75  = quantile(flu_diff_h4, probs = .75,  na.rm = TRUE), #increase
            h4_percentile_95 = quantile(flu_diff_h4, probs = .95, na.rm = TRUE), #large increase
            
            h5_percentile_05 = quantile(flu_diff_h5, probs = .05, na.rm = TRUE), #large decrease
            h5_percentile_25  = quantile(flu_diff_h5, probs = .25,  na.rm = TRUE), #decrease
            h5_percentile_75  = quantile(flu_diff_h5, probs = .75,  na.rm = TRUE), #increase
            h5_percentile_95 = quantile(flu_diff_h5, probs = .95, na.rm = TRUE), #large increase
            
            h6_percentile_05 = quantile(flu_diff_h6, probs = .05, na.rm = TRUE), #large decrease
            h6_percentile_25  = quantile(flu_diff_h6, probs = .25,  na.rm = TRUE), #decrease
            h6_percentile_75  = quantile(flu_diff_h6, probs = .75,  na.rm = TRUE), #increase
            h6_percentile_95 = quantile(flu_diff_h6, probs = .95, na.rm = TRUE), #large increase
            
            h7_percentile_05 = quantile(flu_diff_h7, probs = .05, na.rm = TRUE), #large decrease
            h7_percentile_25  = quantile(flu_diff_h7, probs = .25,  na.rm = TRUE), #decrease
            h7_percentile_75  = quantile(flu_diff_h7, probs = .75,  na.rm = TRUE), #increase
            h7_percentile_95 = quantile(flu_diff_h7, probs = .95, na.rm = TRUE), #large increase
            
            h8_percentile_05 = quantile(flu_diff_h8, probs = .05, na.rm = TRUE), #large decrease
            h8_percentile_25  = quantile(flu_diff_h8, probs = .25,  na.rm = TRUE), #decrease
            h8_percentile_75  = quantile(flu_diff_h8, probs = .75,  na.rm = TRUE), #increase
            h8_percentile_95 = quantile(flu_diff_h8, probs = .95, na.rm = TRUE) #large increase
  ) %>%
  pivot_longer(
    cols = starts_with("h"),
    names_to = c("wk_diff", ".value"),
    names_pattern = "h(\\d)_percentile_(\\d+)"
  ) %>%
  mutate(wk_diff = as.numeric(wk_diff),
         wk_diff = wk_diff +1) %>%
  left_join(most_recent %>%
              select(target_group, value),
            by = c("target_group")) %>%
  mutate(crit1 = `95`,
         crit2 = `75`,
         crit3 = `25`,
         crit4 = `05`) %>%
  select(-c(`95`, `75`, `25`, `05`)) 

```

```{r}
#| label: fig1-heatmap-prep
#| include: false
#| eval: !expr forc_available

# Model names
available_models <- forecast %>% 
  distinct(model) %>%   
  filter(!str_detect(model, "Baseline|Ensemble")) %>% 
  pull()

# convert quantile to categorical
pmf_list <- list()
horizons <- forecast %>% distinct(horizon) %>% pull()
separator <- length(horizons) * (1:(forecast %>% distinct(target_group) %>% pull() %>% length()))
for (model_name in available_models) {
  pmf_list[[model_name]] <- quantile_to_pmf(
    model_id = model_name,
    current_forecasts = forecast,
    threshold = diff_data,
    categories = c("large_increase", "increase", "stable", "decrease", "large_decrease"),
    horizons = horizons
  )
}
output_pmf <- bind_rows(pmf_list)

# ensemble categorical forecast output for this week
task_id_cols <- c("reference_date", "target_group", "horizon", "target_end_date")
categorical_forecast <- output_pmf %>% 
  rename(model_id = model) %>% 
  hubEnsembles::simple_ensemble(
    agg_fun="mean",
    agg_args = list(na.rm = T),
    task_id_cols=task_id_cols
  ) %>% 
  select(-model_id) %>% 
  pivot_wider(names_from = output_type_id,
              values_from = value)

# Split the data by 'group' and create a plot for each subset
fig1_heat <- lapply(unique(categorical_forecast$target_group), function(g) {
  
  heatmap <- cbind(
    raw_truth %>% 
      filter(date < min(forecast$target_end_date)) %>%  
      dplyr::select(target_group, target_end_date = date, value) %>% 
      group_by(target_group) %>% 
      arrange(target_end_date) %>% 
      mutate(flu_diff = value - lag(value, 1)) %>% 
      left_join(diff_data %>% ungroup() %>% filter(wk_diff == 1) %>% select(-c(wk_diff, value)),  # 1-week difference
                by = "target_group") %>% 
      mutate(diff_cat = case_when(flu_diff > crit1 ~ 5, # large increase
                                  flu_diff > crit2 ~ 4, # increase
                                  flu_diff > crit3 ~ 3, # stable
                                  flu_diff > crit4 ~ 2, #decrease
                                  TRUE ~ 1)) %>% 
      select(target_end_date, target_group, diff_cat) %>% 
      filter(target_end_date > as.Date(current_ref) - 150) %>%  
      pivot_wider(names_from = target_end_date   ,
                  values_from = diff_cat ) %>% 
      ungroup() %>% 
      filter(target_group == g) %>% 
      select(-target_group) %>% 
      as.matrix(),
    
    categorical_forecast %>% 
      mutate(max_value = pmax(large_increase, increase, stable, decrease, large_decrease),
             diff_cat = case_when(
               large_increase == max_value ~ 5, # large increase
               increase == max_value ~ 4, # increase
               stable == max_value ~ 3, # stable
               decrease == max_value ~ 2, # decrease
               large_decrease == max_value ~ 1 # large decrease
             )) %>% 
      select(target_end_date, target_group, diff_cat) %>% 
      pivot_wider(names_from = target_end_date,
                  values_from = diff_cat ) %>% 
      ungroup() %>% 
      filter(target_group == g) %>% 
      select(-target_group) %>% 
      as.matrix()
  ) 
  
  # if some week missing
  full_weeks <- seq(from = as.Date(min(colnames(heatmap))), 
                    to = as.Date(max(colnames(heatmap))), 
                    by = "1 week")  
  filled_mat <- matrix(NA, nrow = 1, ncol = length(full_weeks))
  colnames(filled_mat) <- as.character(full_weeks)
  matching_indices <- match(colnames(heatmap), as.character(full_weeks))
  filled_mat[1, matching_indices] <- heatmap
  heatmap <- filled_mat

  hover_text <- cbind(
    raw_truth %>% 
      filter(date < min(forecast$target_end_date)) %>%  
      dplyr::select(target_group, target_end_date = date, value) %>% 
      group_by(target_group) %>% 
      arrange(target_end_date) %>% 
      mutate(flu_diff = value - lag(value, 1)) %>% 
      left_join(diff_data %>% ungroup() %>% filter(wk_diff == 1) %>% select(-c(wk_diff, value)),  # 1-week difference
                by = "target_group") %>% 
      mutate(diff_cat = case_when(flu_diff > crit1 ~ paste0("Week", epiweek(target_end_date), ": ", "Large increase"),
                                  flu_diff > crit2 ~ paste0("Week", epiweek(target_end_date), ": ", "Increase"), 
                                  flu_diff > crit3 ~ paste0("Week", epiweek(target_end_date), ": ", "Stable"), 
                                  flu_diff > crit4 ~ paste0("Week", epiweek(target_end_date), ": ", "Decrease"), 
                                  TRUE ~ paste0("Week", epiweek(target_end_date), ": ", "Large decrease"))) %>% 
      select(target_end_date, target_group, diff_cat) %>% 
      filter(target_end_date > as.Date(current_ref) - 150) %>%  
      pivot_wider(names_from = target_end_date   ,
                  values_from = diff_cat ) %>% 
      ungroup() %>% 
      filter(target_group == g) %>% 
      select(-target_group) %>% 
      as.matrix(),
    
    categorical_forecast %>% 
      mutate(hover_label = paste0("Week", epiweek(target_end_date), 
                                  "<br>Large increase: ", round(large_increase*100,1), "%", 
                                  "<br>Increase: ", round(increase*100,1), "%", 
                                  "<br>stable: ", round(stable*100,1), "%", 
                                  "<br>Decrease: ", round(decrease*100,1), "%", 
                                  "<br>Large decrease: ", round(large_decrease*100,1), "%")) %>% 
      select(target_end_date, target_group, hover_label) %>% 
      pivot_wider(names_from = target_end_date,
                  values_from = hover_label ) %>% 
      ungroup() %>% 
      filter(target_group == g) %>% 
      select(-target_group) %>% 
      as.matrix()
    )
  
  # if some week missing
  full_weeks <- seq(from = as.Date(min(colnames(hover_text))), 
                    to = as.Date(max(colnames(hover_text))), 
                    by = "1 week")  
  filled_mat <- matrix(NA, nrow = 1, ncol = length(full_weeks))
  colnames(filled_mat) <- as.character(full_weeks)
  matching_indices <- match(colnames(hover_text), as.character(full_weeks))
  filled_mat[1, matching_indices] <- hover_text
  hover_text <- filled_mat

  color_scale <- list(
    list(0.0, '#006166'),
    list(0.25, "#3bbbb0"),
    list(0.5, "#e3e3e3"),
    list(0.75, "#b83d93"),
    list(1.0, "#6d085a")
  )
  
  border_shape <- list(
    type = "rect",
    x0 = as.Date(current_ref)+(7*(categorical_forecast %>% distinct(horizon) %>% slice(1) %>% pull()))-3.5,        
    x1 = as.Date(current_ref)+(7*(categorical_forecast %>% distinct(horizon) %>% slice(4) %>% pull()))+3.5,       
    y0 = 0,    
    y1 = 1,    
    line = list(color = "black", width = 3),  
    xref = "x",  
    yref = "paper", 
    fillcolor = "rgba(0,0,0,0)" 
  )
  
  p <- plot_ly(
    x = colnames(heatmap),  
    y = g,  
    z = heatmap,     
    type = "heatmap",
    colorscale = color_scale,
    showscale = FALSE,
    xgap = 1,
    ygap = 1,
    zmin = 1,
    zmax = 5,
    text = hover_text,
    hoverinfo = 'text'
    ) %>% 
    layout(
      yaxis = list(
        showline = FALSE,
        zeroline = FALSE,
        showticklabels = FALSE,
        ticks = "",
        title = NULL,
        showgrid = FALSE,
        fixedrange = TRUE
      ),
      xaxis = list(
        title = "",
        type = "date",
        tick0 = min(fig1_dat$target_end_date),
        dtick = two_weeks_ms,
        range = c(as.Date(min(fig1_dat$target_end_date))-5, as.Date(max(fig1_dat$target_end_date))+5),
        tickvals = tickvals,        # Set tick values to match the target_end_date
        ticktext = ticktext,        # Set tick text to show epiweek and year format
        tickangle = -45
      ),
      shapes = list(border_shape),
      margin = list(t = 50, r = 50)
    ) 
  
  return(p)

})

legend_labels <- c("Large increase", "Increase", "Stable", "Decrease", "Large decrease")
legend_colors <- c("#6d085a", "#b83d93", "#e3e3e3", "#3bbbb0", "#006166")

heatmap_legend_annotations <- lapply(seq_along(legend_labels), function(i) {
  list(
    x = 1,                # just outside right edge
    y = -0.12 + 0.04 * (length(legend_labels) - i),  # space vertically
    xref = "paper",
    yref = "paper",
    showarrow = FALSE,
    text = paste0("<span style='color:", legend_colors[i], "'>\u25A0</span> ", legend_labels[i]),
    font = list(size = 12, family = "Arial"),
    align = "left",
    xanchor = "left"
  )
})
  
```

```{r}
#| label: fig-1-message
#| results: "asis"
#| eval: !expr isFALSE(forc_available)

cat("
<div style='
  display: flex;
  align-items: center;
  justify-content: center;
  height: 200px;
  border: 1px solid #dee2e6;
  border-radius: 6px;
  background-color: #f8f9fa;
  color: #6c757d;
  font-family: Arial, sans-serif;
  font-size: 16px;
'>
  &#9432;&nbsp; This week&#39;s forecast is not available as data was not updated this week.
</div>
")

```


:::{.panel-tabset}
## Overall 
```{r}
#| label: fig-1-plotly-overall
#| results: "asis"
#| warning: false
#| eval: !expr forc_available

plotly::subplot(
  fig1[[2]], 
  fig1_heat[[2]], 
  nrows = 2, 
  heights = c(0.9,  0.1), 
  shareX = FALSE) %>% 
  layout(
    yaxis2 = list(title = list(
                            text = "SARI hospitalizations", 
                            font = list(family="Arial", size = 18),
                            standoff = 15
                            ),
                  automargin = TRUE
                  ),
    annotations = heatmap_legend_annotations,
    legend = list(
      x = 0.01,  
      y = 0.99,  
      tracegroupgap = 1,
      orientation = "v", 
      font = list(family="Arial", size = 12)  
    ),
    margin = list(r = 120)
  )
```

## Adults 
```{r}
#| label: fig-1-plotly-adults
#| results: "asis"
#| warning: false
#| eval: !expr forc_available

plotly::subplot(
  fig1[[1]], 
  fig1_heat[[1]], 
  nrows = 2, 
  heights = c(0.9,  0.1), 
  shareX = FALSE) %>% 
  layout(
    yaxis2 = list(title = list(
                            text = "SARI hospitalizations", 
                            font = list(family="Arial", size = 18),
                            standoff = 15
                            ),
                  automargin = TRUE
                  ),
    annotations = heatmap_legend_annotations,
    legend = list(
      x = 0.01,  
      y = 0.99, 
      orientation = "v",  
      font = list(family="Arial", size = 12)  
    ),
    margin = list(r = 120)
  )

```

## Pediatrics 
```{r}
#| label: fig-1-plotly-pediatrics
#| results: "asis"
#| warning: false
#| eval: !expr forc_available

plotly::subplot(
  fig1[[3]], 
  fig1_heat[[3]], 
  nrows = 2, 
  heights = c(0.9,  0.1), 
  shareX = FALSE) %>% 
  layout(
    yaxis2 = list(title = list(
                            text = "SARI hospitalizations", 
                            font = list(family="Arial", size = 18),
                            standoff = 15
                            ),
                  automargin = TRUE
                  ),
    annotations = heatmap_legend_annotations,
    legend = list(
      x = 0.01,  
      y = 0.99,  
      orientation = "v",  
      font = list(family="Arial", size = 12) 
    ),
    margin = list(r = 120)
  )

```
:::

## Prior weeks forecast performance

To evaluate the retrospective performance of prior weeks’ hindcast/nowcast/forecasts, we overlay the truth data (i.e., observed SARI cases) from the current week, depicted with the black dots, with the forecast outputs (i.e., predicted SARI cases). By comparing the observed data to what was expected, we can evaluate the performance of our forecast models. The red dots correspond to the data that were available at the time the forecast models were ran each week.

```{r}
#| label: retrospective-helper
#| include: false

generate_forecast_plots <- function(current_week, current_forecast) {
  # Calculate the number of weeks between the current week and the first forecast week
  num_weeks <- as.integer(difftime(current_week, first_forecast, units = "weeks"))
  
  # load time-series (append version)
  training_data <- read_csv(
       paste0(here("target-data"),"/time-series.csv"),
        col_types = cols_only(
          as_of = col_character(),
          target_end_date = col_character(),
          target = col_character(),
          age_group = col_character(),
          observation = col_double())
       ) %>%
      mutate(as_of = as.Date(parse_date_time(as_of, orders = c("Ymd", "mdy", "mdY"))),
             target_end_date = as.Date(parse_date_time(target_end_date, orders = c("Ymd", "mdy", "mdY"))))

  plots_list <- list()

  # Loop through the weeks and generate the forecast plots
  for (wk in 1:num_weeks) {
    # Date setting
    current_ref  <- current_week - lubridate::days(wk * 7)
    epi_week <- lubridate::epiweek(current_ref)
    
    forecast <- tryCatch({
      # Ensemble forecast output for the given week
      readr::read_csv(
        paste0(here("auxiliary-data", "microhub-output"), 
               "/microhub-output_",current_ref,".csv"),
         col_types = cols_only(
            model = col_character(),
            reference_date = col_character(),
            horizon = col_integer(),
            target_end_date = col_character(),
            target_group = col_character(),
            output_type = col_character(),
            output_type_id = col_character(),
            value = col_double()
          )
        ) %>%
      mutate(reference_date = as.Date(parse_date_time(reference_date, orders = c("Ymd", "mdy", "mdY"))),
             target_end_date = as.Date(parse_date_time(target_end_date, orders = c("Ymd", "mdy", "mdY")))) %>% 
      filter(output_type == "quantile") 
    }, error = function(e) NULL)
    
    # Skip this week if file was not found or failed to load
    if (is.null(forecast)) next
    
    # pull corresponding training data
    week_start <- current_ref - 6  # Sunday
    week_end   <- current_ref      # Saturday
    archive_dates <- training_data %>% dplyr::distinct(as_of) %>% pull()
    target_archive <- archive_dates[archive_dates >= week_start & archive_dates <= week_end]
    if (length(target_archive) != 1) {
      stop(sprintf(
        "Expected 1 archive file for week of %s, found %d.", current_ref, length(target_archive)
      ))
    }

    forc_data <- training_data %>%
      filter(as_of == target_archive) %>% 
      select(date = target_end_date, target_group = age_group, value = observation) %>% 
      arrange(date) 

    # Plotly data to work with
    common_dates <- intersect(forc_data %>% pull(date), 
                              forecast$target_end_date)
    feb_1 <- lubridate::make_date(year = lubridate::year(current_ref), month = 2, day = 1)
    nov_1 <- lubridate::make_date(year = lubridate::year(current_ref), month = 11, day = 1)
    fig2_dat <- raw_truth %>%
      filter(date < min(current_forecast$target_end_date)) %>%
      dplyr::select(target_end_date = date, target_group, value) %>% 
      full_join(bind_rows(forc_data %>% 
                            filter(date < min(forecast$target_end_date)) %>% 
                            dplyr::select(target_end_date = date, target_group, used_value = value),
                          forecast %>%
                            filter(model == "Ensemble") %>%
                            pivot_wider(names_from = output_type_id,
                                        values_from = value) %>% 
                            dplyr::select(forc_value = `0.5`, target_end_date, target_group,
                                          `0.025`, `0.975`, `0.25`, `0.75`)),
                by = c("target_group", "target_end_date")) %>%
      filter(target_end_date > feb_1) %>%
      filter(target_end_date < nov_1) %>%
      group_by(target_group) %>% 
      complete(target_end_date = seq(min(target_end_date), max(target_end_date), by = "1 week")) %>% 
      mutate(
        hindcast_drop = case_when(target_end_date %in% common_dates ~ forc_value, TRUE ~ NA_real_),
        hindcast = case_when(!(target_end_date %in% common_dates) & 
                               target_end_date < current_ref ~ forc_value, TRUE ~ NA_real_),
        nowcast  = case_when(target_end_date == current_ref ~ forc_value, TRUE ~ NA_real_),
        forecast = case_when(target_end_date > current_ref ~ forc_value, TRUE ~ NA_real_)
      ) %>% 
      mutate(
        week_label = if_else(
          epiweek(target_end_date) == 1,
          paste0(lubridate::epiyear(target_end_date), " Week ", epiweek(target_end_date)),
          paste0("Week ", epiweek(target_end_date)))
      ) 

    # Create a list of unique dates and their corresponding epiweek labels for tickvals and ticktext
    tickvals <- seq.Date(min(fig2_dat$target_end_date), nov_1, by = "1 week")
    ticktext <- if_else(
        lubridate::epiweek(tickvals) == 1,
        paste0(lubridate::epiyear(tickvals), " Week ", lubridate::epiweek(tickvals)),
        paste0("Week ", lubridate::epiweek(tickvals))
      )

    # x axis tick
    two_weeks_ms <- 14 * 24 * 60 * 60 * 1000

    # Split the data by 'group' and create a plot for each subset
    plots <- lapply(unique(fig2_dat$target_group), function(g) {
      
      ymax <- global_ymax[[g]]

      dat <- fig2_dat %>% filter(target_group == g) %>% 
        mutate(
          purple_line = coalesce(hindcast_drop, hindcast, nowcast),
          blue_line = coalesce(nowcast, forecast)
        )
      
      p <- plot_ly(data = dat,
                   x = ~target_end_date) %>%
        add_trace(y = ~`0.025`, name = "outer_lower",
                  type = "scatter", mode = "lines", line = list(color = "transparent"),
                  showlegend = FALSE,
                  yaxis = 'y2') %>%
        add_trace(y = ~`0.975`, name = "outer_upper",
                  type = "scatter", mode = "lines", line = list(color = "transparent"),
                  fill = "tonexty", fillcolor = "#CBEAFB",
                  showlegend = FALSE,
                  yaxis = 'y2') %>%
        add_trace(y = ~`0.25`, name = "inner_lower",
                  type = "scatter", mode = "lines", line = list(color = "transparent"),
                  showlegend = FALSE,
                  yaxis = 'y2') %>%
        add_trace(y = ~`0.75`, name = "inner_up",
                  type = "scatter", mode = "lines", line = list(color = "transparent"),
                  fill = "tonexty", fillcolor = "#5CC3F0",
                  showlegend = FALSE,
                  yaxis = 'y2') %>%
        add_trace(y = ~value, name = "Reported",
                  type = "scatter", mode = "lines+markers",
                  line = list(color = "black"),
                  marker = list(color = "black", size = 8),
                  yaxis = 'y2') %>%
        add_trace(y = ~used_value, name = "Training",
                  type = "scatter", mode = "markers",
                  marker = list(color = "darkred", size = 8),
                  yaxis = 'y2') %>%
        add_trace(y = ~purple_line, name = "purple_line",
                  type = "scatter", mode = "lines",
                  line = list(color = "#7B6FC4"),
                  showlegend = FALSE, yaxis = 'y2') %>%
        add_trace(y = ~blue_line, name = "blue_line",
                  type = "scatter", mode = "lines",
                  line = list(color = "#2B55D4"),
                  showlegend = FALSE, yaxis = 'y2') %>%
        add_trace(y = ~hindcast_drop, name = "Hindcast (dropped weeks)",
                type = "scatter", mode = "markers",
                # line = list(color = "#7B6FC4"),
                marker = list(color = "transparent", size = 8, line = list(color = "#7B6FC4", width = 2)),
                showlegend = FALSE,
                yaxis = 'y2') %>%
        add_trace(y = ~hindcast, name = "Hindcast",
                  type = "scatter", mode = "markers",
                  # line = list(color = "#7B6FC4"),
                  marker = list(color = "#7B6FC4", size = 8),
                  yaxis = 'y2') %>%
        add_trace(y = ~nowcast, name = "Nowcast",
                  type = "scatter", mode = "markers",
                  # line = list(color = "#1A3A8F"),
                  marker = list(color = "#1A3A8F", size = 8),
                  yaxis = 'y2') %>%
        add_trace(y = ~forecast, name = "Forecast",
                  type = "scatter", mode = "markers",
                  # line = list(color = "#2B55D4"),
                  marker = list(color = "#2B55D4", size = 8),
                  yaxis = 'y2') %>% 
        add_trace(y = ~value, "Y1-Forecast",
                  type = "scatter", mode = "lines",
                  line = list(color = "transparent"),
                  showlegend = FALSE,
                  yaxis = 'y1') %>%
        layout(
          title = list(text = "",
                       x = 0.05, font = list(family = "Arial", size = 18, style = "bold")),
          xaxis = list(title = "",
                       type = "date",
                       tick0 = min(fig2_dat$target_end_date),
                       dtick = two_weeks_ms,
                       range = c(min(tickvals), max(tickvals)),
                       tickvals = tickvals,       # Set tick values to match the target_end_date
                       ticktext = ticktext,        # Set tick text to show epiweek and year format
                       tickangle = -45
                        ),
          yaxis = list(title = "",
                       showticklabels = FALSE, showgrid = FALSE,
                       range = c(0, ymax)),
          yaxis2 = list(title = list( text = "SARI hospitalizations",
                                      font = list(family="Arial", size = 18),
                                      standoff = 15 ),
                        side = "right", showline = TRUE, showticklabels = TRUE,
                        tickangle = 0, tickfont = list(size = 15),
                        automargin = TRUE,
                        range = c(0, ymax)),
          legend = list(x = 0.01, y = 0.99, orientation = "v", font = list(family = "Arial", size = 12)),
          annotations = list(
            list(x = 0.99, y = 0.98, xref = "paper", yref = "paper",
                 text = "Inner Bands: 50% Prediction Intervals", font = list(family = "Arial", size = 12),
                 showarrow = FALSE),
            list(x = 0.99, y = 0.93, xref = "paper", yref = "paper",
                 text = "Outer Bands: 95% Prediction Intervals", font = list(family = "Arial", size = 12),
                 showarrow = FALSE)
          ),
          margin = list(t = 50, r = 50)
        )

      return(p)
    })


    # Store the plots and the corresponding epi week
    plots_list[[wk]] <- list(epi_week = epi_week, plots = plots)
  }

  # # Reverse the order of the plots_list
  # plots_list <- rev(plots_list)

  return(plots_list)
}


```

```{r}
#| label: retrospective-plots
#| message: false
#| include: false

# Pick the most recent forecast
most_recent <- forc_files[which.max(as.Date(forc_dates))]

# Ensemble forecast output
forecast_recent <- readr::read_csv(
  most_recent,
   col_types = cols_only(
      model = col_character(),
      reference_date = col_character(),
      horizon = col_integer(),
      target_end_date = col_character(),
      target_group = col_character(),
      output_type = col_character(),
      output_type_id = col_character(),
      value = col_double()
    )
  ) %>%
mutate(reference_date = as.Date(parse_date_time(reference_date, orders = c("Ymd", "mdy", "mdY"))),
       target_end_date = as.Date(parse_date_time(target_end_date, orders = c("Ymd", "mdy", "mdY")))) %>% 
filter(output_type == "quantile") 

# Compute global y-axis max across all forecast submissions
global_ymax <- forc_files[grepl(epiyear(current_ref), forc_files) & forc_files != most_recent] %>%
  purrr::map(function(f) {
    tryCatch({
      readr::read_csv(f,
        col_types = cols_only(
          model          = col_character(),
          output_type    = col_character(),
          output_type_id = col_character(),
          target_group   = col_character(),
          value          = col_double()
        )) %>%
        filter(model == "Ensemble",
               output_type == "quantile",
               output_type_id == "0.975") %>%
        group_by(target_group) %>%
        summarise(max_val = max(value, na.rm = TRUE), .groups = "drop")
    }, error = function(e) NULL)
  }) %>%                     
  dplyr::bind_rows() %>% 
  group_by(target_group) %>%
  summarise(global_ymax = max(max_val, na.rm = TRUE) * 1.3, .groups = "drop") %>% 
  tibble::deframe()       


# Generate forecast plots
plots_list <- tryCatch({
    generate_forecast_plots(current_week = current_ref, current_forecast = forecast_recent)
}, error = function(e) {
    NULL
})

```
:::{.panel-tabset}
```{r retro-forecast-tabs, echo=FALSE, results='asis', fig.show='hold'}
# setting
plots_list <- compact(plots_list)
n_weeks <- length(plots_list)
n_show  <- seq_len(n_weeks)
plot_order <- c(2, 1, 3)
tab_labels <- c("Overall", "Adults", "Pediatrics")

# for each week
for (i in n_show) {
  wk <- plots_list[[i]]
  week <- wk$epi_week[1]

  cat(paste0("## Week ", week, "\n\n"))
  
  # open nested tabset
  cat("::: {.panel-tabset}\n\n")

  for (k in seq_along(plot_order)) {
    j <- plot_order[k]
    cat(paste0("### ", tab_labels[k], "\n\n"))
    html_snip <- renderTags(wk$plots[[j]])$html
    cat(html_snip, "\n\n")
  }
  
  # close nested tabset
  cat(":::\n\n")
}

```
:::


## Weighted Interval Score (WIS) table

The Weighted Interval Score (WIS) table displays the interval scores for each forecast model. Forecasts with lower absolute WIS values are considered more accurate than forecasts with higher absolute WIS values. The relative WIS compares each forecast model to the regular baseline model; the regular baseline model simply predicts that influenza activity in the next five weeks will be a straight line with the band of uncertainty expanding each week. Mean absolute error (MAE) values are also reported which is a measure of forecast accuracy, or how well the forecast models are performing. A lower MAE indicates a more accurate forecast.

```{r}
#| label: wis-data
#| include: false

# define weeks we want to evaluate
last_forcasting <- current_ref - lubridate::days(7)
weeks_to_eval <-
  seq(first_forecast, last_forcasting, by=7) %>%
  as.character()

# Forecast output
raw_forecasts <- tryCatch({
  list.files(
    path    = here("auxiliary-data", "microhub-output"),
    pattern = paste0("microhub-output_(", paste(format(as.Date(weeks_to_eval), "%Y-%m-%d"), collapse = "|"), ")\\.csv$"),
    full.names = TRUE
  ) %>%
  lapply(read.csv) %>%
  do.call(rbind, .) %>% 
  mutate(reference_date = as.Date(reference_date),
         target_end_date = as.Date(target_end_date))
}, error = function(e) NULL)

# data to calculate WIS
raw_data_wis <- tryCatch({
  raw_forecasts %>%
    left_join(
      raw_truth %>%
        filter(date < min(forecast$target_end_date)) %>% 
        dplyr::select(target_end_date = date, target_group, observed = value) ,
      by = c("target_group", "target_end_date")
    ) %>%
    rename(quantile_level=output_type_id, predicted=value) %>%
    mutate(quantile_level = as.numeric(quantile_level))
}, error = function(e) NULL)

# calculate scores (used for WIS and other metrics)
# Error message is expected
raw_scores <- tryCatch({
  raw_quantile <- raw_data_wis %>%
    select(-output_type) %>% 
    scoringutils::as_forecast_quantile(
      forecast_unit = c("model", "reference_date", "horizon", "target_end_date", "target_group")
    ) 
  metrics <- get_metrics(raw_quantile)
  metrics$interval_coverage_95 <- purrr::partial(interval_coverage, interval_range = 95)
  raw_quantile %>% 
    scoringutils::score(metrics = metrics)
}, error = function(e) NULL)


```

```{r}
#| label: wis-message
#| results: "asis"
#| eval: !expr is.null(raw_scores)

cat("
<div style='
  display: flex;
  align-items: center;
  justify-content: center;
  height: 200px;
  border: 1px solid #dee2e6;
  border-radius: 6px;
  background-color: #f8f9fa;
  color: #6c757d;
  font-family: Arial, sans-serif;
  font-size: 16px;
'>
  &#9432;&nbsp; Since observed case counts for the prior forecasting horizon have not yet been recorded or reported, scoring cannot be performed at this time.
</div>
")

```

```{r}
#| label: tabset-open
#| results: "asis"
#| eval: !expr "!is.null(raw_scores)"

cat(":::{.panel-tabset}\n")
```

```{r}
#| label: tab1-header
#| results: "asis"
#| eval: !expr "!is.null(raw_scores)"

cat("### 6-week WIS\n\n")
```

```{r}
#| label: wis-6wk
#| eval: !expr "!is.null(raw_scores)"

# Helper function to calculate wis 
compute_wis <- function(scores, filter_expr = NULL, log_scale = FALSE) {
  if (!is.null(filter_expr)) scores <- scores %>% filter(!!filter_expr)
  if (log_scale) {
    scores <- scores %>%
      mutate(wis = ifelse(wis == 0, 1e-05, wis),
             wis = log(wis))
  }
  scores %>%
    add_relative_skill(compare = "model", baseline = "Regular Baseline") %>%
    summarise_scores(by = "model") %>%
    mutate(wis     = round(wis, 2),
           rel_wis = round(wis_scaled_relative_skill, 2),
           mae     = round(ae_median, 2),
           cov_50  = round(interval_coverage_50 * 100, 2),
           cov_95  = round(interval_coverage_95 * 100, 2)) %>%
    select(model, wis, rel_wis, mae, cov_50, cov_95)
}

# Count number of forecasts evaluated (6 week)
n_6wk_date <- raw_scores %>%
  filter(reference_date >= last_forcasting - (7*5)) %>%
  summarise_scores(by = c("model", "target_group","reference_date")) %>%
  mutate(baseline_n_forecasts = sum(!is.na(wis[model == "Regular Baseline"]))) %>%
  group_by(model) %>%
  mutate(n_forecasts = sum(!is.na(wis))) %>%
  mutate(per_forecasts = round((n_forecasts/(baseline_n_forecasts))*100,2))

### WIS, log-WIS, MAE, and coverage (6-week)
filter_6wk <- quote(reference_date >= last_forcasting - (7 * 5))
wis_6wk <- compute_wis(raw_scores, filter_6wk) %>% 
  left_join(compute_wis(raw_scores, filter_6wk, log_scale = TRUE) %>% 
              select(model, log_wis = wis, rel_log_wis = rel_wis),
            by = "model")

# WIS table (6-week)
n_6wk_date %>%
  select(model, n_forecasts, per_forecasts) %>%
  distinct() %>%
  left_join(wis_6wk,
            by="model") %>%
   rename(Model = model,
          `Absolute WIS` = wis,
          `Relative WIS`= rel_wis,
          `Absolute log WIS` = log_wis,
          `Relative log WIS`= rel_log_wis,
          `50% Coverage (%)` = cov_50,
          `95% Coverage (%)` = cov_95,
          `Number of Forecasts Submitted`  = n_forecasts,
          `% Forecasts Submitted` = per_forecasts,
          `MAE` = mae) %>%
  as.data.frame() %>%
  arrange(`Relative WIS`, `Absolute WIS`) %>%
  select(Model, `Absolute WIS`, `Relative WIS`, `Absolute log WIS`, `Relative log WIS`,
         MAE, `50% Coverage (%)`,
         `95% Coverage (%)`,
         `Number of Forecasts Submitted`,
         `% Forecasts Submitted`) %>%
  datatable(options = list(dom = 't'))

```

```{r}
#| label: tab2-header
#| results: "asis"
#| eval: !expr "!is.null(raw_scores)"

cat("### Season WIS\n\n")
```

```{r}
#| label: wis-season
#| eval: !expr "!is.null(raw_scores)"

# Count number of forecasts evaluated (season)
n_season_date <- raw_scores %>%
  summarise_scores(by = c("model", "target_group","reference_date")) %>%
  mutate(baseline_n_forecasts = sum(!is.na(wis[model == "Regular Baseline"]))) %>%
  group_by(model) %>%
  mutate(n_forecasts = sum(!is.na(wis))) %>%
  mutate(per_forecasts = round((n_forecasts/(baseline_n_forecasts))*100,2))

### WIS, MAE, and coverage (season)
filter_season <- quote(epiyear(reference_date) >= epiyear(last_forcasting))
wis_season <- compute_wis(raw_scores, filter_season) %>% 
  left_join(compute_wis(raw_scores, filter_season, log_scale = TRUE) %>% 
              select(model, log_wis = wis, rel_log_wis = rel_wis),
            by = "model")

# WIS table
n_season_date %>%
  select(model, n_forecasts, per_forecasts) %>%
  distinct() %>%
  left_join(wis_season,
            by="model") %>%
   rename(Model = model,
          `Absolute WIS` = wis,
          `Relative WIS`= rel_wis,
          `Absolute log WIS` = log_wis,
          `Relative log WIS`= rel_log_wis,
          `50% Coverage (%)` = cov_50,
          `95% Coverage (%)` = cov_95,
          `Number of Forecasts Submitted`  = n_forecasts,
          `% Forecasts Submitted` = per_forecasts,
          MAE = mae) %>%
  as.data.frame() %>%
  arrange(`Relative WIS`, `Absolute WIS`) %>%
  select(Model, `Absolute WIS`, `Relative WIS`, `Absolute log WIS`, `Relative log WIS`,
         MAE, `50% Coverage (%)`,
         `95% Coverage (%)`,
         `Number of Forecasts Submitted`,
         `% Forecasts Submitted`) %>%
  datatable(options = list(dom = 't'))

```

```{r}
#| label: tabset-close
#| results: "asis"
#| eval: !expr "!is.null(raw_scores)"

cat(":::\n")
```

## Weekly reported influenza cases

This interactive chart shows how the influenza case data
evolve and stabilize over time as new data are submitted to FluNET. Because
FluNET data can be revised or backfilled after initial reporting, this plot is
designed to visualize both the availability of data for forecasting and the 
extent to which these data stabilize over time. Hovering over
any element will display the corresponding details.

The <u>colored dots</u> represent successive data submissions. The color 
indicates how many weeks after the first reportable week each submission was 
made - that is, weeks since data for that week could first be complete. 
The yellow dot marks a submission made in the first reportable week (0 weeks),
while darker pink/purple dots indicate later submissions as
countries update or backfill their data. The maximum number of submissions since 
first reportable week is capped at five in this chart, 
as meaningful data revisions beyond that point are uncommon and reduce visual 
clarity when graphed. To ensure that submissions with identical values remain 
individually visible rather than collapsing into a single dot, the orange dots 
are horizontally jittered by a small random offset.

The <u>black line</u> connects the most recently reported data for each epidemiological
week and represents the most up-to-date influenza case counts. When the black
dot appears close to earlier yellow/orange dots, the data are relatively 
stable and required little to no revision.

```{r}
#| label: stab-data
#| include: false

# load time-series (append version)
training_data <- read_csv(
     paste0(here("target-data"),"/time-series.csv"),
      col_types = cols_only(
        as_of = col_character(),
        target_end_date = col_character(),
        target = col_character(),
        age_group = col_character(),
        observation = col_double())
     ) %>%
    mutate(as_of = as.Date(parse_date_time(as_of, orders = c("Ymd", "mdy", "mdY"))),
           target_end_date = as.Date(parse_date_time(target_end_date, orders = c("Ymd", "mdy", "mdY"))))

# manipulate
currentyear <- training_data %>% 
  group_by(as_of, age_group, target_end_date) %>%
    summarise(
      value = sum(observation),
      .groups = "drop"
    ) %>%
  arrange(as_of, age_group, target_end_date) %>% 
  group_by(as_of) %>%
  mutate(admit_week = epiweek(target_end_date),
         publish_year = epiyear(as_of),
         publish_week = epiweek(as_of)) %>%
  select(target_group = age_group, 
         date = target_end_date, admit_week, publish_year, publish_week, publish_date = as_of,
         value) %>% 
  filter(epiyear(date) >= epiyear(current_ref)) %>% 
  arrange(target_group, date, publish_date) %>% 
  group_by(target_group, date) %>% 
  mutate(week_since = (publish_week - 1) - admit_week)

currentyear_last <- currentyear %>% 
  slice_tail(n = 1) %>%      
  ungroup() %>%
  arrange(admit_week)

target_week <- epiweek(current_ref) - 1
currentyear_first <- currentyear %>% 
  slice_head(n = 1) %>%      
  group_by(target_group) %>% 
  complete(admit_week = full_seq(c(admit_week, target_week), 1)) %>%
  ungroup() %>%
  arrange(target_group, admit_week)

set.seed(42)
currentyear <- currentyear %>%
  arrange(desc(week_since)) %>% 
  filter(week_since <= 5) %>% 
  mutate(x_jitter = admit_week + runif(n(), min = -0.1, max = 0.1))

```

```{r}
#| label: stab-graph-helper
#| include: false

make_stab_plot <- function(group, currentyear, currentyear_last, currentyear_first) {
  
  range <- currentyear %>%
    filter(target_group == group) %>%
    group_by(admit_week) %>%
    summarise(y_min = min(value, na.rm = TRUE),
              y_max = max(value, na.rm = TRUE), .groups = "drop")

  finalized_label <- paste0(
    "Week ",
    unique(currentyear_last %>% filter(target_group == group) %>% pull(publish_week)),
    " data"
  )
  
  unique_weeks <- currentyear_first %>% 
    filter(target_group == group) %>% 
    pull(admit_week) %>% 
    unique() %>% 
    sort()

  plot_ly(data = currentyear %>% filter(target_group == group), x = ~x_jitter) %>%
    add_trace(
      y = ~value, name = "Submission dots", type = "scatter", mode = "markers",
      text = ~paste0("Week ", admit_week, ": ", value, " (available in week ", publish_week, ")"),
      hovertemplate = "%{text}<extra></extra>", showlegend = FALSE,
      marker = list(
        color = ~week_since,
        cmin = 0,
        cmax = 5,
        colorscale = list(
          list(0.0, "#FFB74D"), list(0.2, "#EF6C00"), list(0.4, "#B71C1C"),
          list(0.6, "#E91E8C"), list(0.8, "#7B1FA2"), list(1.0, "#4A148C")
        ),
        size = 10, opacity = 0.8,
        colorbar = list(title = "Weeks since<br>first reportable week", tickfont = list(family = "Arial", size = 12),
                        tickmode = "array", tickvals = list(0,1,2,3,4,5),
                        ticktext = list("0","1","2","3","4","5"))
      )
    ) %>%
    add_segments(
      data = range, x = ~admit_week, xend = ~admit_week,
      y = ~y_min, yend = ~y_max,
      line = list(color = "#FF6B00", width = 0.5), hoverinfo = "skip", showlegend = FALSE
    ) %>%
    add_trace(
      data = currentyear_last %>% filter(target_group == group),
      x = ~admit_week, 
      y = ~value, 
      name = finalized_label, 
      type = "scatter", mode = "lines+markers",
      text = ~paste0("Current value: ", value), 
      hovertemplate = "%{text}<extra></extra>",
      line = list(color = "black", width = 1.5), 
      marker = list(color = "black", size = 10),
      showlegend = TRUE
    ) %>%
    layout(
      title  = list(text = "", x = 0.05, font = list(family = "Arial", size = 25)),
      xaxis = list(
        title = "",
        range = c(min(unique_weeks) - 0.5, max(unique_weeks) + 0.5),
        tickvals = unique_weeks, ticktext = paste("Week", unique_weeks),
        tickfont = list(family = "Arial", size = 12), tickangle = -45,
        showgrid = TRUE, gridcolor = "lightgrey", gridwidth = 1
      ),
      yaxis  = list(title = "Number of SARI hospitalizations", showline = TRUE,
                    tickfont = list(family = "Arial", size = 14),
                    range = c(0, max(currentyear$value, na.rm = TRUE) * 1.3)),
      showlegend = TRUE,
      legend = list(x = 0.01, y = 0.99, orientation = "v", traceorder = "reversed",
                    font = list(family = "Arial", size = 12)),
      margin = list(t = 50, r = 50)
    )
}

```

::: {.panel-tabset}
## Overall
```{r}
#| label: stab-graph-1
#| fig.width: 14
#| fig.height: 10

make_stab_plot("Overall", 
               currentyear = currentyear, 
               currentyear_last= currentyear_last, 
               currentyear_first= currentyear_first)

```

## Adults
```{r}
#| label: stab-graph-2
#| fig.width: 14
#| fig.height: 10

make_stab_plot("Adults", 
               currentyear = currentyear, 
               currentyear_last= currentyear_last, 
               currentyear_first= currentyear_first)

```

## Pediatrics
```{r}
#| label: stab-graph-3
#| fig.width: 14
#| fig.height: 10

make_stab_plot("Pediatrics", 
               currentyear = currentyear, 
               currentyear_last= currentyear_last, 
               currentyear_first= currentyear_first)

```
:::

## Appendix 

### Table A1: categorical forecast thresholds

```{r}
#| label: appendix-table
#| results: "asis"

if (!forc_available) {
  
  cat("
<div style='
  display: flex;
  align-items: center;
  justify-content: center;
  height: 200px;
  border: 1px solid #dee2e6;
  border-radius: 6px;
  background-color: #f8f9fa;
  color: #6c757d;
  font-family: Arial, sans-serif;
  font-size: 16px;
'>
  &#9432;&nbsp; This week&#39;s forecast is not available as data was not updated this week.
</div>
")
  
} else {
  
  diff_data %>%
    mutate(order = case_when(target_group == "Overall" ~ 1,
                             target_group == "Adults" ~ 2,
                             target_group == "Pediatrics" ~ 3)) %>% 
    mutate(horizon = wk_diff+(horizons[1]-1)) %>% 
    filter(horizon %in% horizons) %>% 
    arrange(order) %>% 
    ungroup() %>% 
    select(-c(order, value)) %>%
    relocate(target_group, wk_diff, horizon) %>% 
    rename("Age group" = target_group,
           "Horizon" = horizon,
           "Week difference" = wk_diff,
           "95th percentile<br>(large increase)" = crit1,
           "75th percentile<br>(increase)" = crit2,
           "25th percentile<br>(decrease)" = crit3,
           "5th percentile<br>(large  decrease)" = crit4) %>% 
    kable(align = "c", format = 'html', escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>% 
    row_spec(separator,                                                            # Separator between groups
             extra_css = "border-bottom: 1.3px solid black;") %>%
    column_spec(1, bold = TRUE, color = "#000000", width = "3cm") 
  
}

```

